---
title: "R Programming Language"
author: "Jia-Ling Zheng"
output: html_document
---
108-1 NCKU Stat上課筆記整理

# R
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
* [Basic]
  + [Assignment]
  + [Calculation]
  + [Sequence]
  + [Naming]
  + [Approximated values]
* [Vector]
  + [Numeric Vector]
  + [Character vectors]
  + [Logical Vector]
* [Matrix]
  + [Construct a matrix]
  + [Extract elements from a matrix]
* [ControlFlow]
  + [for() loop]
  + [if() statement]
  + [while() loop]
  + [break, repeat, and next]
  + [Some Examples]
* [Functions]
  + [source( )]
  + [function( ) with one single input]
  + [function( ) with multiple input variables]
  + [function( ) and return multiple output]
* [DataFrame]
  + [Data Frame]
  + [Data input and output at hard disk]
  + [R Package]
  + [Tools for help]
* [Graphics]
  + [Statistical Figures]
  + [High-levlel plotting functions]
  + [Low-levlel plotting functions]
* [RMarkdown]
* [Apply family]
  + [Function: apply(. . . )]
  + [Function: lapply(. . . )]
  + [Function: sapply(. . . )]
* [Aboutdata]
  + [Remove NA]
  + [Data Processing]
  + [Package: reshape2]
  + [Regression]
  + [Package: dplyr]
  
## Basic
### Assignment

We recommended using “<-” because it makes clear
that we are requesting an assignment, while "="
is stating a relation.

### Calculation

There are still many arithmetic operators we don't mention. 

[Search for “Arithmetic Operators” in R.](https://www.datamentor.io/r-programming/operator/)
```{r}
5 + 3 ### Addition
5 * 3 ### Multiplication
5 / 3 ### Division
5 ^ 3 ### Exponent
5 %% 3 ### Modulus(Remainder from division)
5 %/% 3 ### Integer Division
NA  ### Missing value
NaN ### 0/0 (Not a Number)
Inf ### 2/0 (Infinity)
exp(2) ; log(10)### Exponential;logarithm
```
### Sequence
more explanation : type "?seq"
```{r}
1:10
seq(1, 10, 2)
seq(1, 10, length.out = 6)
```
### Naming
  1. Avoid function name or the specific words
     + break; while; Inf; NaN 
  2. Recommended Naming a meaningful object by "." or "_"
     + height_Amy, weight_old
   
### Approximated values
```{r}
round(125.78253, digits = 3); round(125.78253, digits = -2);
signif(83524587, digits = 3); signif(83524587, digits = 5);
floor(125.78253) ###無條件捨去
ceiling(125.78253) ###四捨五入
trunc(125.78253) ### extract the integer part
```
## Vector
### Numeric Vector
* Three ways to create a vector 
```{r}
vt1 <- c(1:5)  
vt1
seq(1, 5, by = 0.5) ; seq(1,15, length.out=9)
rep(c(3,9),each=2) ; rep(seq(2, 10, 2), 2)

```
* Creat a null or initial vector

  + c( ): with unknown the length.
  
  + rep(NA, n): With known length.
```{r}
a <- c(3, 5, 9, 10)
b <- c(14, 15, 17)
V3 <- rep(NA, 7) 
V3[1:4] <- a
V3
V3[5:7] <- b
V3
```
* Elements in the vector
  + unique( ) : return the elements that appear at least one time.
  + table( )  : count the number of each element from the vector.
```{r}
b <- c(3, 3, 3, 3, 5, 5, 1, 2, 10, 10, 8)
unique(b)
table(b)
```
* Set Operations
```{r}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
union(B, C)
intersect(B, C)
setdiff(B, C)
setequal(B, C)
```
* Sorting and ordering of a vector
  + sort(x, decreasing = FALSE, . . . )
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
sort(a)
sort(a, decreasing = TRUE)
```
* Get the position (index) of sorting of a vector
  + order(. . . , na.last = TRUE, decreasing = FALSE)
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
inc.order <- order(a)
dec.order <- order(a, decreasing = TRUE)
inc.order
a[inc.order]
dec.order
a[dec.order]
```
* Change values in a vector
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
a[5] <- -3
a[8] <- -1
a
```
* Deleting elements from a vector
```{r}
vt1
vt1[-2]
vt1[-c(2:3)]
```
### Character vectors
* Print Characters
```{r}
print("Hello R World!!!")
```
* Split a string
* Lowercase and uppercase
  + toupper(): change to uppercase
  + tolower(): change to lowercase
```{r}
TEXT1 <- "Let's go to a game."
strsplit(TEXT1, " ")
toupper(TEXT1)
tolower(TEXT1)
```
* Character vectors

- the vector is made up of strings of characters.

  "All elements of a vector must be of the same type."
```{r}
colors <- c("red", "yellow", "BLUE")
more.colors <- c(colors, "green", "cyan")
more.colors
z <- c("red", "green", 12) ###attempt to mix data types in a vector
z
```
* Basic operations for character vectors

  substr( ) or grep() or paste()
  
  1. To take substrings, use “substr( )”.
```{r}  
colors <- c("red", "yellow", "blue")
substr(colors, 1, 3)
```
  2. To search a string from a vector, use “grep()”, which is “Global Regular
Expression Print”.

It returns the locations of the desired patterns.
```{r}
colors <- c("red", "yellow", "BLUE")
more.colors <- c(colors, "green", "cyan")
grep("y", more.colors)
#How to show the whole character from grep()?
y.in.V <- grep("y", more.colors)
more.colors[y.in.V]
```
  3. To building up strings, use “paste()”.
```{r}
colors <- c("red", "yellow", "blue")
paste(colors, "line") ## with space
paste(colors, "line", sep = "" ) ## no space
paste("I like", colors, collapse = "; ")
```
* Factors

factor(): factor(x = character())
```{r}
colors <- c("red", "yellow", "blue")
f.c <- factor(colors)
f.c
levels(f.c)
as.integer(f.c)
```
### Logical Vector
* Logical number

  -Binary operators which allow the comparison of values in atomic vectors (relational operators).
```{r}
c(3 < 5, 3 > 5)
c(3 <= 5, 3 >= 5)
3 == 5 ###equal
3 != 5 ###not equal
```
* Boolean operations
```{r}
A <- c(T, T, F, T, F); B <- c(3, 1, 5, 4, 7)
A
B[A]
!A
B[!A]
A <- c(T, T, F, T, F)
```

* Logical vector

  -Compare the same location of elements individually.
```{r}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
B == C
D <- c(2, 5 ,6, 10)
B == D
```
* “is.” series
  -“is.XYZ()” is any comparison to the type “XYZ” of an object.

     is.null() ; is.na() ; is.nan() ; is.infinite()

     is.character() ; is.integer() ; is.numeric() ;
    
```{r}
is.null(c())
c(is.na(NA), is.na(2))
x1 <- 1:5
x2 <- c(1:3, 2.2)
x3 <- colors
c(is.integer(x1), is.numeric(x1), is.double(x1))
c(is.integer(x2), is.numeric(x2), is.double(x2))
c(is.character(x3), is.numeric(x3))
```
## Matrix
### Construct a matrix
  1. Construct a matrix by row
  2. Construct a matrix by column
```{r}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20
M1 <- rbind(a, b, c, d)
M1
M2 <- cbind(a, b, c, d)
M2
```
  3. Construct a matrix by a long vector
```{r}
#預設是由上而下由右至左排序
matrix(1:20, nrow = 4, ncol = 5)
matrix(1:20, nrow = 4, ncol = 5, byrow = TRUE)
```
* Names in a matrix
```{r}
M1 <- matrix(1:20, nrow = 4, ncol = 5, byrow = TRUE)
dim(M1)
colnames(M1) <- c("V1", "V2", "V3", "V4", "V5")
dimnames(M1)
```
### Extract elements from a matrix
  - We still use “[ ]” to extract elements from a matrix, but
  
     using a “,” to identify the locations of row and column seperately.
```{r}
M1 <- matrix(1:20, nrow = 4, ncol = 5, byrow = TRUE)
M1[c(2, 4), c(3, 5)];
#Extract a row or a column from a matrix
M1[, 4]; M1[3, ]
colnames(M1) <- c("V1", "V2", "V3", "V4", "V5")
rownames(M1) <- c("a", "b", "c", "d")
rownames(M1)
dim(M1)
#Extract elements by names
############M1[c("a","d"),c("v2", "v3") ]
```
### Simple evaluation on matrix
```{r}
A <- matrix(c(1:4), 2, 2)
B <- matrix(c(5:8), 2, 2)
A + 3
A - 1
A * 5
A / 4
A * B
```
* Matrix multiplication

  - Use %*% to evaluate multiplication of matrices.
```{r}
A <- matrix(c(1:4), 2, 2)
B <- matrix(c(5:8), 2, 2)
A %*% B
```
* About Matrix 
```{r}
B <- matrix(c(5:8), 2, 2)
t(B) ### Transpose
diag(B) ### Diagonal matrix
det(B) ### Determinant of a matrix
solve(B) ### Inverse matrix
eigen(B) ### Eigen system
```
* Summary each column or each row
```{r}
C <- matrix(1:9, 3, 3)
colSums(C)
rowSums(C)
colMeans(C)
rowMeans(C)
```

## ControlFlow
### for() loop
  - specify that a certain operation be repeated a fixed number of times.
  
**Syntax**

              for (index in vector) commands,
  
where

+ index: the index is changed each time
 
+ vector: it includes all possible cases

+ commands: actions

### if() statement
- One condition:
  if (condition) {
  
    commands when TRUE
    
    } else {
    
    commands when FALSE
    
    }
- Multiple conditions:
  if (condition1) {
  
    commands when condition1 is TRUE
    
    } else if (condistion2) {
    
    commands when condition2 is TRUE
    
    } else {
    
    commands when condition1 and condition2 are FALSE
    
    }

### while() loop
  - The pattern of repetition isn’t known in advance and we need to do some calculations and keep going as long as a condition holds. 

**Syntax**

              while (condition) statements
If the condition evaluates to FALSE, nothing more is done. If it evaluates to
TRUE the statements are executed, condition is evaluated again, and the
process is repeated.

+ Relationship between for() and while()
  - for()
```{r}
for(j in 1:8){
  print(j)
}
```
  - while(): Need the initial, condition of stopping, and iteration or replacement
```{r}
j <- 1
while(j <= 8){
  print(j)
  j <- j + 1
}
```

### break, repeat, and next
+ break

  if (condition) break

  The break statement causes the loop to terminate immediately. break statements
  can also be used in for() and while() loops.

+ repeat

  Use “repeat” to repeat until we execute a break statement.

  repeat { statements }
  
  This causes the statements to be repeated endlessly.

+ next

  The next statement causes control to return immediately to the top of the loop; it can also be used in any loop.
  
### Some Examples

+ Flip a coin

  Let 0 be a tail and 1 be a head. Count the total number of flipping until each success occurs.
```{r}
coin <- c(0, 0, 0, 0, 0, 0, 1, 0, 1, 0, 1, 0,
          1, 1, 0, 1, 1, 0, 0, 0, 1, 0, 1, 0,
          0, 0, 0, 0, 0, 0, 1)
N <- c(); j <- 1
while(j <= length(coin)){
  n.fail <- 0
  while(coin[j] != 1){
    n.fail <- n.fail + 1
    j <- j + 1
    }
  N <- c(N, n.fail + 1)
  j <- j + 1
}
N
```
+ Fibonacci numbers

  List all Fibonacci numbers less than 100. Use for() and if(condition) break.
```{r}
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
for(j in 1:1000){
  if(Fib2 < 100){
    Fibonacci <- c(Fibonacci, Fib2)
    Fib1 <- Fibonacci[length(Fibonacci) - 1]
    Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
  }else{
    break
  }
}
Fibonacci
```

## Functions
### source( )
1. Creat a new script for your function or codes.
2. **Save** the script named as “XXX.R” and use **source( )** to run the code.
3. Remember to use **print( )** or **cat( )** so that your final answer will be
shown.
+ print( ) & cat( )
  - **print( )** shows with " " between characters.
  - **cat()** shows without " " between characters. You can add ”\n" to make characters in a new line.
```{r}
x <- c(0.8933, 0.2344, 0.0764, 0.7736)
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
print(x.show)
cat(x.show)
```
```{r}
x <- c("A", "B", "C", "D")
print(x)
cat(x)
cat(x[1], "\n",
x[2], "\n", sep = "" )
```

### function( ) with one single input

Use **function( )** when you only have one main code for evaluation and
multiple possible input, like $$f (x) = 3x + 2, x = 2, 3, 4.$$
f means a name of a function of input x when x = 2, 3, 4.
```{r}
f <- function(x){
  3*x + 2
}
f(2); f(3); f(4)
```
+ return( ) in the function( )

You must tell what will be return when you use the function by **return( )** or **invisible( )**.

If you don’t put **return( )** at the end of the function, it will treat the last line as the result automatically.
```{R}
f <- function(x){
return(3*x + 2)
}
f(2)
```
```{R}
f <- function(x){
res <- 3*x + 2
}
f(2) ## return nothing
a <- f(2); a
```

```{R}
make.percent <- function(x){
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
return(x.show) 
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans <- make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans
```
```{r}
make.percent <- function(x){
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
invisible(x.show) ## compare to return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans <- make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
ans
```

### function( ) with multiple input variables
$$f (x, y) = x^2 + y^2, x, y = 2, 3, 4.$$
f means a name of a function of input x and y when x, y = 2, 3, 4.
```{r}
f <- function(x, y){
x^2 + y^2
}
f(2, 2); f(2, 3); f(2, 4)
```
+ Example: percent
```{r}
make.percent <- function(x, x.digits){
x.percent <- round(100*x, digits = x.digits)
x.show <- paste(x.percent, "%", sep = "")
return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 0)
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 2)
```
+ Set a default in the function

Add the settings of default in the arguments of input variables.

Example: percent
```{r}
make.percent <- function(x, x.digits = 0){
x.percent <- round(100*x, digits = x.digits)
x.show <- paste(x.percent, "%", sep = "")
return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 0)
```

### function( ) and return multiple output
+ Introuction to list( )
**list( )** is a collection of multiple types of objects, like vectors, characters, numbers, lists,. . . .

Example:
```{r}
X <- list(a = 3, b = c(1,2,3,4,5), c = c("A", "B"))
X
```
+ Extract an element from a list

Use **$** or **[[ ]]**.

  - **$**: the name of the element.
  - **[[ ]]**: the location or the name of the element.
```{r}
X <- list(a = 3, b = c(1,2,3,4,5), c = c("A", "B"))
X$a
X[[2]]
X[["c"]]
```
+ Creat a list
```{r}
try.list <- list() ## a null list
length(try.list)
```
```{r}
length(try.list) <- 2 ## assign the size to the list
try.list[[1]] <- seq(0, 10, length.out = 3) ## without name
try.list$square <- seq(0, 10, length.out = 3)^2 ## with name
try.list
```
+ Exact element from the list
```{r}
try.list[[3]]
try.list$square
names(try.list)
length(try.list)
```

+ Use function( ) and list( )

Example: perfect number
```{r}
perfect.number <- function(a, b){
  k <- a
  res <- c()
  ans <- list()
  while(k < b){
    fac <- c()
    for(j in 1:k){
      if((k %% j) == 0) fac <- c(fac, j)
    }
    if(k == sum(fac[-length(fac)])) {
      res <- c(res, k)
      ans[[paste(k)]] <- fac[-length(fac)]
    }
      k <- k+1
  }
  ans$number <- res
  return(ans)
}
Ans <- perfect.number(1, 100)
Ans
```

## DataFrame
### Data Frame
  1. Most data sets are stored in R as data frames.
  2. These are like matrices, but with the columns having their own names.
  3. Columns can be of different types from each other.
```{r}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers,
more.numbers = c(4, 5, 6))
colors.and.numbers
```
+ Difference from matrix
  1. These are like matrices, but with the columns having their own names.
  2. Columns can be of different types from each other.
```{r}
X <- matrix(NA, 3, 3)
X[,1] <- c("red", "yellow", "blue")
X[,2] <- c(1, 2, 3)
X[,3] <- c(4, 5, 6)
colnames(X) <- c("colors", "numbers", "more.numbers")
X
```
+ Types in a data frame: character

Change the type from factor to character by using stringAsFactors = FALSE.
```{r}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers,
more.numbers = c(4, 5, 6),
stringsAsFactors = FALSE)
str(colors.and.numbers)
```
+ More on data.frame(): add rows
  Add column and row vectors into a data frame.
  - row: use rbind()
  - column: use the sign $ or cbind()
  
  Use **stringsAsFactors = FALSE**
```{r}
data2 <- data.frame(colors, numbers, more.numbers = c(4, 5, 6),
stringsAsFactors = FALSE)
data2 <- rbind(data2, c("black", 3, 5))
data2
```
+ More on data.frame(): add columns
  Add column and row vectors into a data frame.
  - column: use the sign $ or cbind()
  - row: use rbind()
```{r}
colors <- c("red", "yellow", "blue")
number1 <- c(1, 2, 3)
number2 <- c(10, 11, 12)
colors.and.numbers <- data.frame(colors, number1) ## construct colors.and.numbers$more.numbers <- c(4, 5, 6) ## add one column
colors.and.numbers <- cbind(colors.and.numbers, number2) ## add colors.and.numbers
```

### Data input and output at hard disk
1. Change and find directory
  - Change the directory: setwd(“dir”)
  - Find the directory: get()
```{r eval=FALSE}
setwd("C:/Users/stat/Desktop")
getwd()
```
2. Save and load the objects 

Suppose you have constructed some R objects, type **dump(objects, file = “XXX.R”)**.
```{r}
X <- seq(0, 10, 0.1)
A <- 3
B <- mean(X)
dump(list = objects(), file = "all.R")
dump(c("A", "X"), file = "parts.R")
```
To retrieve the objects in a future session, type **source(“XXX.R”)**
```{r}
source("all.R")
source("parts.R")
```
+ To save all objects in the current workspace, type **save.image(“XXX.RData”)**.
+ To save parts of objects, type **save(a, b, c, file = “XXX.RData”)**.
+ Use **load(“XXX.RData”)** to retrieve data.
```{r eval=FALSE}
X <- seq(0, 10, 0.1)
A <- 3
B <- mean(X)
save.image("all.RData")
save(A, B, "parts.RData")
load("parts.RData")
load("all.RData")
```

3. Save data to a file

You can save a matrix, a vector, or a data frame into two types of file, .txt (.dat) and .csv.

+ Use **write.table(A, file = “XXX.txt”)**
+ Use **write.csv(A, file = “XXX.csv”)**
```{r}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
X <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
write.table(X, "matrix.txt")
write.csv(X, "matrix.csv")
```

4. Read data from a file 

You can laod a matrix, a vector, or a data frame from two types of file, .txt
(.dat) and .csv.

+ Use **read.table(file = “XXX.txt”)**
+ Use **read.csv(file = “XXX.csv”)**
```{r eval=FALSE}
X1 <- read.table("matrix.txt")
X2 <- read.csv("matrix.csv")
```
With names of columns in a .txt file, you should add **header = TRUE**.
```{r eval=FALSE}
read.table("test.txt", header = TRUE)
```

### R Package
1. Introduction to R packages
+ Other than some basic functions, some developed methodologies and
tools are provided in a way of packages. You can download them from
CRAN or Github.
+ To download a package, type install.packages(“XXX”).
+ Before using some functions in a downloaded package, you should call
the package by typing library(XXX).

```{r eval=FALSE}
install.packages("data.table")
```
2. Use the functions in the installed packages
+ Before using some functions in a downloaded package, you should call
the package by typing library(XXX).
Try the function fread( ), which is similar to read.table( ) but faster and
more convenient.
```{r eval=FALSE}
library(data.table)
mydat <- fread('test.csv')
head(mydat)
```
3. Read data though internet
```{r eval=FALSE}
input <- fread("http://www.stats.ox.ac.uk/pub/datasets/csb/ch11b.dat")
head(input)
```
4. More on packages

[CRAN - Contributed Packages](https://cran.r-project.org/web/packages/)

[R bloggers](https://www.r-bloggers.com/october-2019-top-40-new-r-packages/)

### Tools for help
1. Online help
```{r eval=FALSE}
help(mean)
?median
```
2. Built-in examples
```{r}
example(mean)
```
3. Only knowing a keyword of function
```{r eval=FALSE}
help.search("optimization")
```

## Graphics
### Statistical Figures

1. Descriptive Statistics 

  Given an arbitray vector x:
  
  + Histogram: hist(x)
  + Boxplot: boxplot(x)
  + Density plot: plot(density(x))

```{r}
x.norm <- rnorm(100, mean=0, sd=1)
hist(x.norm)
boxplot(x.norm)
plot(density(x.norm))
```
    
  For discrete variables x:
  
  - Barchart: barplot(x)
  - Piechart: pie(x)

```{r}
x <- sample(c(1:8), 100, replace=TRUE, prob=c(8:1))
x <- table(x)
names(x) <- c("A", "B", "C", "D", "E", "F", "G", "H")
barplot(x)
pie(x)
```

  For multiple categories:
  
  + Scatter plot
  + Boxplot
```{r}
x <- rnorm(100)
y <- 3*x + rnorm(100, 0, 1)
plot(x, y)
cor(x, y)
X <- matrix(rnorm(30), 10, 3) ## use matrix
pairs(X, pch = 19)
## use matrix
X <- matrix(rnorm(50), 10, 5) 
boxplot(X)
## use data.frame
X <- data.frame(group = rep(c("A", "B", "C", "D", "E"), each = 10),
res = rnorm(50))
boxplot(res ~ group, data = X)
```
  
### High-levlel plotting functions
  + plot(. . . )

        plot(x, y,
      
            type = "p", ### types of the plot
      
            pch = , ### the symbol in the plot
       
            main = NULL, ### the title of the plot
      
            lty = , ### the style of a line
       
            lwd = , ### the width of a line
      
            cex = , ### the size of an text
      
            bty="o", ### the box around the plot
      
            col = , ### the color of an object
      
            xlim = NULL, ylim = NULL, # the range of an axis
      
            sub = NULL, ### the sub-title of the plot
      
            xlab = NULL, ylab = NULL, # the name of an axis
      
            cex.main = ### the size of the title
        )
  1. plot(. . . , type = . . . )
      
```{r}
par(mfrow = c(2,3)) # split plot ranges
plot(x, y, type = "p", main = "p") #points
plot(x, y, type = "l", main = "l") #lines
plot(x, y, type = "b", main = "b") #both points and lines
plot(x, y, type = "o", main = "o") #lines pass through points
plot(x, y, type = "h", main = "h") #the lines between points and x-axis
plot(x, y, type = "n", main = "n") #no content in the plot but plot region
```
 
  2. plot(. . . , pch = . . . )
  
      Choose a symbols for points.
      
      pch = c(1, 2, ..., 25, "*", ".", "o", "O","0", "+", "-", "|", "%", "#")
      
  3. plot(. . . , cex = . . . )
  
      Sizes in a plot: the number of times size of the default one
      - Size of points: cex = 1.5
      - Size of the title: cex.main = 2
      - Size of the labels of ticks of an axis: cex.axis = 2
      - Size of the labels of the name of an axis: cex.lab = 2   
```{r}
par(mfrow = c(2,3)) # split plot ranges
plot(x, y, pch = 1, main = "pch = 1")
plot(x, y, pch = 3, main = "pch = 3", cex.main = 2)
plot(x, y, pch = 10, main = "pch = 10", cex.axis = 2)
plot(x, y, pch = 15, main = "pch = 15", cex = 2, cex.main = 2)
plot(x, y, pch = 19, main = "pch = 19", cex = 2, cex.axis = 2)
plot(x, y, pch = "#", main = "pch = #", cex = 2,
cex.axis = 2, cex.lab = 2, cex.main = 2)
```

  4. plot(. . . , type = “l”, lty =. . . , lwd = . . . )
    - The style of lines: lty = 1, 2, . . . , 7
    - The width of lines: lwd = 1, 1.25,. . .
```{r}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2,2)) # split plot ranges
plot(t, y, type = "l")
plot(t, y, type = "l", lty = 2, lwd = 2)
plot(t, y, type = "l", lty = 3, lwd = 2)
plot(t, y, type = "l", lty = 5, lwd = 2)
```

  5. plot(. . . , bty = . . . )
  
    The box around the plot:
  
    bty = c("o", "n", "7", "L", "C", "U")
```{r}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2,3)) # split plot ranges
plot(t, y, type = "l", bty = "o", main = "bty = \" o \"") #o: complete box (default parameter)
plot(t, y, type = "l", bty = "n", main = "bty = \" n \"") #n: no box
plot(t, y, type = "l", bty = "7", main = "bty = \" 7 \"") #7: top & right
plot(t, y, type = "l", bty = "L", main = "bty = \" L \"") #L: bottom & left
plot(t, y, type = "l", bty = "C", main = "bty = \" C \"") #C: top & left & bottom
plot(t, y, type = "l", bty = "U", main = "bty = \" U \"") #U: left & bottom & right
```

  6. plot(. . . , col = . . . )
    - Use number: col = c(1, 2, 3, 4, 5, 6, 7, 8)
    - Use name: col = c(“red”, “green”, “blue”, “grey”)
    - Use hexadecimal code: col = c(“#009999”, “#0000FF”, . . . )
    - More on colors: help(palette)
```{r}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2,3)) # split plot ranges
plot(t, y, type = "l", col = 1)
plot(t, y, type = "l", col = 2)
plot(t, y, type = "l", col = "green")
plot(t, y, type = "l", col = "grey")
plot(t, y, type = "l", col = "#009999")
plot(t, y, type = "l", col = "#0000FF")
```
  
  7. plot(. . . , main = . . . , sub = . . . , xlab = . . . , ylab = . . . )
    - main: main title (at the top)
    - sub: sub title (at the bottom)
    - xlab: name of x-axis
    - ylab: name of y-axis
```{r}  
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2,2)) # split plot ranges
plot(t, y, type = "l")
plot(t, y, type = "l", main = "Time series")
plot(t, y, type = "l", main = "Time series", sub = "this year")
plot(t, y, type = "l", main = "Time series", xlab = "time", ylab = "return")
```
  
  8. plot(. . . , xlim = . . . , ylim = . . . )
    
    Ranges of variables
```{r}
t <- 1:20
y <- rnorm(20)
par(mfrow = c(2,2)) # split plot ranges
plot(t, y, type = "b")
plot(t, y, type = "b", xlim = c(0, 50))
plot(t, y, type = "b", ylim = c(-5, 5))
plot(t, y, type = "b", xlim = c(0, 50), ylim = c(-5, 5))
```
  
  + curve(. . . )
  
  Instead using plot(x, y, type = “l”), use curve(expression, from = ..., to = ...)
  to draw a function of a variable.
```{r}  
par(mfrow = c(2, 3))
curve(exp(-x), 0, 10)
curve(dnorm(x), -10, 10)
curve(x^2, -10, 10)
curve(sin, 0, 10)
curve(3*exp(-3*x), 0, 6)
curve(5*x+2, 0, 10)
```

### Low-levlel plotting functions
  + Add points or lines to an existing plot
    - points(. . . ): add points
    - lines(. . . ): add lines
    - abline(. . . ): add straight lines
    
    abline(a = ..., b = ..., h = ..., v = ..., reg = ..., coef = ...)
    $$y = a + bx$$

```{r}
plot(t, y, type = "b")
abline(h = 0)
abline(v = 15, lty = 2, col = 8, lwd = 2)
points(t[15], y[15], pch = 19, col = 2)
```

  + Add text to an existing plot
    - Arbitrary locations:
  
      text(x, y, labels = . . . )
    - Notations of objects in the plot
    
      legend(x, y, legend, . . . )
```{r}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)
points(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)
text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1,4), lty = c(1, 1),
cex = 1.2)
```

  + Re-arrange the region of a figure
  
     Use par(. . . ) to creat an arbitrary region
    - mfrow = c(2, 2): Define a multi-figure display by row
    - mfcol = c(2, 2): Define a multi-figure display by column
    - mai, mar = c(3, 3, 1, 1): The four margin sizes, in inches (mai), or in text line units (mar).
    
```{r}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
par(mfrow = c(1, 1), mar = c(0.5, 0.5, 0.5, 0.5))
plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)
points(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)
text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1,4), lty = c(1, 1),
cex = 2)
```
  
```{r}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
y2 <- -2*x + rnorm(50, 0, 2)
layout(matrix(c(1,2,3,3), 2, 2, byrow=TRUE))
plot(x, y, type = "p", pch = 19, main = "Y")
abline(a = 0, b = 3)
plot(x, y1, type = "p", pch = 19, main = "Y1")
abline(a = 0, b = 5)
plot(x, y2, type = "p", pch = 19, main = "Y2")
abline(a = 0, b = -2)
```

  + Split the whole plot by some locations of plot
    - layout(matrix(c(1,2,3,3), 2, 2, byrow=TRUE))

## RMarkdown
  + Compile a single R Markdown document to a report in different
    formats, such as PDF, HTML, or Word.
  + Create notebooks in which you can directly run code chunks
    interactively.
  + Make slides for presentations (HTML5, LaTeX Beamer, or PowerPoint).

  1. The first step:
```{r eval=FALSE}
install.packages("rmarkdown")
```
  2. Create a new file
  
File → New File → Rmarkdown → Choose Title/Author/Output format(HTML/PDF/WORD) → OK

### Some basic editings
    *italics* and _italics_
    **bold** and __bold__
    superscript^2^ and ~~strikethrough~~
  *italics* and _italics_
  
  **bold** and __bold__
  
  superscript^2^ and ~~strikethrough~~
  
### Headers of a Paragraph

  - Use “#” to show the paragraph.
  
  *#* Header 1
    
  *##* Header 2
    
  *###* Header 3
    
  *####* Header 4
    
  *#####* Header 5
      
# Header 1
## Header 2
### Header 3
#### Header 4
##### Header 5

### Equations
    $A = \pi*r^{2}$
    $$B = 2 \pi r$$
Inline equation:

$A = \pi*r^{2}$

Center an eqiation:

$$B = 2 \pi r$$

### Bullet point

    * unordered list
    * item 2
        + sub-item 1
    1. ordered list
    2. item 2
        + sub-item 1
* unordered list
* item 2
    + sub-item 1
1. ordered list
2. item 2
    + sub-item 1

### Table 

    Table Header  | Second Header
    ------------- | -------------
    Table Cell    | Cell 2
    Cell 3        | Cell 4

Table Header  | Second Header
------------- | -------------
Table Cell    | Cell 2
Cell 3        | Cell 4

    \begin{table}[h]
        \centering
        \begin{tabular}{c|cc}
            \hline
            $x_j$ & 0 & 1 \\
            \hline
            $p_j$ & 0.3 & 0.7 \\
            \hline
        \end{tabular}
        \caption{Random variable $X$ with its probability}
    \end{table}

\begin{table}[h]
    \centering
    \begin{tabular}{c|cc}
        \hline
        $x_j$ & 0 & 1 \\
        \hline
        $p_j$ & 0.3 & 0.7 \\
        \hline
    \end{tabular}
    \caption{Random variable $X$ with its probability}
\end{table}

### Matrix

    $\left[
        \begin{array}{cc}
          A & B \\
          C & D \\
        \end{array}
      \right]$
  
$\left[
    \begin{array}{cc}
      A & B \\
      C & D \\
    \end{array}
  \right]$

  
### Sources:

  - [Introduction to R Markdown](https://rmarkdown.rstudio.com/lesson-1.html)
  - [Beamer presentation](https://bookdown.org/yihui/rmarkdown/beamer-presentation.html)
  - [Beamer-theme-matrix](https://hartwork.org/beamer-theme-matrix/)
  - [Chunk options and package options](https://yihui.org/knitr/options/)
  - [The R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

## Apply family
### Function: apply(. . . )
  **apply(X, MARGIN, FUN, . . . )** Returns a vector or array or list of values obtained by applying a function to margins of an array or a matrix.
  
   - X: array or matrix (or a data frame)
   - MARGIN: for a matrix 1 indicates rows, 2 indicates columns
   - FUN: name of a function
```{r results='hide'}
A <- matrix(rnorm(1500), 500, 3)
#Sum of each row
apply(A, 1, sum); rowSums(A)
#Sum of each column
apply(A, 2, sum); colSums(A)
```

Write a function for the sum of each row.   
```{r}
row.col.sum <- function(X, margin = 1){
  size <- dim(X)
  if(margin == 1){
    res <- rep(NA, size[1])
    for(j in 1:size[1]){
      res <- c(res, sum(X[j,]))
    }
  }else if(margin == 2){
    res <- c()
    for(j in 1:size[2]){
      res <- c(res, sum(X[,j]))
    }
  }else{
    stop("margin should be either 1 or 2.")
  }
  return(res)
}
```

+ Check time-efficiency
```{r}
library(rbenchmark)
benchmark(replications=rep(1000, 1),
        apply(A, 1, sum), rowSums(A), row.col.sum(A, 1),
        columns=c('test', 'elapsed', 'relative'))
```

+ What if there are NA’s?
```{R}
B <- matrix(sample(1:100, 15), 5, 3)
B[1, 2] <- NA
B[3, 1] <- NA
B <- data.frame(B)
colnames(B) <- c("factor 1", "factor 2", "factor 3")
apply(B, 2, mean)
```

Recall: mean(. . . , na.rm = TRUE)
```{R}
apply(B, 2, mean, na.rm = TRUE)
```

+ Other function in apply(X, MARGIN, FUN)
```{r}
apply(B, 2, max)
apply(B, 2, min, na.rm = TRUE)
```

### Function: lapply(. . . )
  **lapply(X, FUN, . . . )** Returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.

  - X: vector or data frame or list
  - FUN: name of a function
```{r}
C <- list(a = 1:10, beta = exp(-3:3))
lapply(C, mean)
par(mfrow = c(1, 2))
lapply(C, hist)
```
  
### Function: sapply(. . . )
  **sapply(X, FUN, . . . , simplify = TRUE)** Similarlly to lapply(X, Fun, . . . ) if sapply(x, f, simplify = FALSE, USE.NAMES = FALSE)

  - X: vector or data frame or list
  - FUN: name of a function 
  - simplify: logical; TRUE will return an array or a matrix with the same length of output
```{r}
sapply(C, summary)
sapply(C, summary, simplify = FALSE)
D <- list(a = sample(1:10, 10, replace = TRUE),
b = sample(1:5, 20, replace = TRUE))
sapply(D, unique)
sapply(D, table)
```

## Aboutdata
### Remove NA
+ If NA, . . .
  - complete.cases(X): X could be a sequence of vectors, matrices and data frames.
    - na.omit(X)
```{r}
dim(airquality)
no.na <- complete.cases(airquality)
head(airquality[no.na,])
dat.no.na.1 <- airquality[no.na,]
dim(dat.no.na.1)
dat.no.na.2 <- na.omit(airquality)
dim(dat.no.na.2)
```
### Data Processing
+ Multiples datasets
  - Using **cbind(. . . )** and **rbind(. . . )** to combine datasets.
  - Using **cut(. . . )** to group or to label items.
  
For example, would like to lable 5 groups by the variable “Wind”.
```{r}
grouped1 <- cut(airquality[,3], 5)
head(grouped1)
grouped2 <- cut(airquality[,3], 5,
labels = c("G1", "G2", "G3", "G4", "G5"))
head(grouped2)
```
+ Summary the grouped data
1. Summarize the number of each group.
```{r}
airquality$group <- grouped2
table(grouped2)
```
+ Odering the dataset
2. After labeling the group, order the dataset by the group. **order(. . . )**
```{r}
order.data.1 <- airquality[order(airquality$group),]
head(order.data.1)
```
If there is no column for further sorting, it returns the result sorted by the index of
the row.

3. Keep ordering the dataset by the temprature. **order(first, second, . . . )**
```{r}
order.data.2 <- airquality[order(airquality$group, airquality$Temp), ]
head(order.data.2)
```

### Package: reshape2
+ Change the format from original one

Package: “reshape2”

For example, the dataset is as follows:
```{r}
game <- c("G1", "G2", "G3", "G4", "G5")
site <- c("M", "O", "L", "O", "L")
Lin <- c(15, 6, 26, 22, 18)
Jordan <- c(18, 32, 21, 25, 12)
Peter <- c(10, 6, 22, 9, 12)
ballgames <- data.frame(game, site, Lin, Jordan, Peter)
```
```{r eval=FALSE}
install.packages("reshape2")
```
+ melt(. . . )

Use **melt(data, id.vars)** to reshape the dataset.
```{r}
library(reshape2)
after.reshape <- melt(ballgames, id.vars = c("game", "site"))
```
+ dcast(. . . )

Use **dcast(dataset, formula, fun)** to show the specific data format.
```{r}
dcast(after.reshape, game + site ~ variable, sum)
dcast(after.reshape, variable ~ game, sum)
dcast(after.reshape, variable ~ site, mean)
dcast(after.reshape, variable + site ~ game, sum)
```
+ Merge data

Two datasets:
```{r}
Lin.game <- ballgames[ballgames[,"Lin"] > 20,
c("game", "site", "Lin")]
Jordan.game <- ballgames[ballgames[,"Jordan"] > 20,
c("game", "site", "Jordan")]
Lin.game
Jordan.game
```
Using **merge(. . . )** to combine datasets to a new dataset by

1. intersection
2. union: set **all = TURE** in merge(. . . )
```{R}
merge(Lin.game, Jordan.game)
merge(Lin.game, Jordan.game, all = TRUE)
```

Merge data by the information of the specific dataset.

+ merge(x, y, all.x = TRUE)
+ merge(x, y, all.y = TRUE)
```{r}
merge(Lin.game, Jordan.game, all.x = TRUE)
merge(Lin.game, Jordan.game, all.y = TRUE)
```

### Regression
+ linear model: lm(. . . )
  - Regression is used to show the relationship between independent variables and dependent variables.
  - Before data analysis, we should check the relation plot.
```{r}
data1 <- airquality[1:120, ]
data2 <- airquality[121:153, ]
lm1 <- lm(Ozone ~ Temp, data = data1)
lm2 <- lm(Ozone ~ Temp - 1, data = data1)
summary(lm1)
summary(lm2)
```
+ Prediction by the fitted linear model
```{r}
y.predict <- predict(lm1, data2)
y.predict
```
+ Performance of fitted model
```{r}
plot(airquality[, 4], airquality[, 1], pch = 19)
abline(lm1)
points(airquality[121:153, 4], y.predict, col = 2, pch = 19)
```

+ Performance of prediction
```{r}
plot(airquality[121:153, 1], y.predict, pch = 19)
abline(0, 1)
```

### Package: dplyr
+ %>% in package: dplyr

**%>%** means that apply functions to an object.
```{r eval=FALSE}
library(magrittr)
y <- sample(1:100, 10)
mean(y)
y %>% mean()
min(y)
y %>% min()
```
+ Multiple functions
```{r eval=FALSE}
sample(1:100, 5) %>% mean()
sample(1:100, 10) %>% min()
```
+ Write a code to report the last name of each member.
```{r eval=FALSE}
nba_stars <- c("Stephen Curry", "LeBron James")
names <- strsplit(nba_stars, split = " ")
toupper(names[[1]][2])
nba_stars %>%
  strsplit(, split = " ") %>%
  `[[` (1) %>%
  `[` (2) %>%
  toupper()
```